---
title: "Week10 Assignment"
author: "Mohit Mehndiratta"
date: "10/14/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# DATA PROCESSING AND VISUALISATION

## Question 1

### 1. Import the abalone dataset and explore.

```{r}
library(dplyr)
library("ggplot2")
library(psych)

abalone <- read.csv("../datasets/abalone.csv", stringsAsFactors = TRUE)
head(abalone)
dim(abalone)
summary(abalone)
str(abalone)
```

The abalone data set has 4177 observations of 9 variables. Each variable is numeric in nature except the variable "Sex" which is a factor variable.

### 2. Subset the dataset by selecting the variables Length, Diameter and Height. 

```{r}
new_data <- abalone %>% select(Length, Diameter, Height)
head(new_data)
```

### 3. Subset the dataset by selecting observations of female abalone.

```{r}
female <- abalone %>% filter(Sex == "F")
head(female)
```

### 4. Subset the dataset by selecting the weight variables of Male abalone.

```{r}
male_weight <- abalone %>% filter(Sex == "M") %>% select(Sex, Whole.weight,
                                                         Shucked.weight,
                                                         Viscera.weight,
                                                         Shell.weight)
head(male_weight)
```

### 5. Summarize the mean-length by gender

```{r}
mean_length <- abalone %>% group_by(Sex) %>% summarise(mean = mean(Length))
mean_length
```

## Question 2


### 1. Create a scatter plot of matrices for the dataset.

```{r}
pairs.panels(abalone)
```


### 2. Create a scatter plot of Diameter Vs Length

```{r}
ggplot(data = abalone, mapping = aes(Length, Diameter)) + geom_point(col = "blue")
```

### 3. Create a set of histograms of Shell-weight by Sex category and fill with different colurs. (use facet_wrap with 1 column)

```{r}
ggplot(data=abalone, mapping=aes( x=Shell.weight, fill=Sex)) +
  geom_histogram()+
  facet_wrap(~Sex, ncol=1)
```

### 4. Create violin plot for Shell-weight by Sex category.

```{r}
ggplot(data = abalone, mapping = aes(x = Sex, y = Shell.weight, fill = Sex)) +
  geom_violin()
```

### 5. Create box plot for Shell-weight by Sex category.Enhance the plot by adding title, labels etc. . .

```{r}
ggplot(data = abalone, mapping = aes(x = Sex, y = Shell.weight, fill = Sex)) +
  geom_boxplot() + 
  labs(title = "Boxplots for Sex Categories w.r.t Shell-Weight", x = "Sex"
       , y = "Shell Weight")
```

