---
title: "Week4_Assignment"
author: "Mohit Mehndiratta"
date: "18/08/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1
### (a) - Load and attach the “Default” dataset and view the first few rows. Investigate the structure/class of each variable using str() function.

We will be loading the Default dataset directly from ISLR library.

```{r}
library(ISLR)
attach(Default)
head(Default)
dim(Default)
str(Default)
```

Here we can see that, there are a total of 10,000 observations of 4 variables. Out of which "default" and "student" are categorical/factor variables and "balance", "income" are numerical variables.

Since, default is the categorical variable we can use it in our logistic regression model.

### (b) - Fit a Logistic Regression Model for default in terms of balance and give the model.

```{r}
model1 <- glm(default~balance, data = Default, family = binomial)
summary(model1)
```

The model1 above, is our logistic regression model for default in terms of balance. Upon looking at the summary we can see that value of the **intercept** is **-10.65** and the **slope** is **0.005499**. Therefore our equation for this model will be:

$\hat{P(Y)}$ = $\frac{e^{\hat{\alpha} + \hat{\beta} X}}{1 + e^{\hat{\alpha} + \hat{\beta} X}}$

After replacing the intercept and slope value w.r.t to default vs balance model:

$\hat{P(default)}$ = $\frac{e^{\hat{-10.65} + \hat{0.005499} balance}}{1 + e^{\hat{-10.65} + \hat{0.005499} balance}}$


This is the probability that a person with the balance X would be in default.

### (c) - Give the probability that a person with balance 500 will be in default.

So Given that,

balance = 500, by replacing this in our model equation we can find out the probability.

```{r}
balance = 500
t = -10.65 + (0.005499 * balance) # this is our exponential's superscript.
# probability
probabilityOfDefault = exp(t) / (1 + exp(t))
probabilityOfDefault
```

From the output, we can see that **0.000370421** is the probability of a person with balance 500 to be in a default.


### (d) - Fit a Logistic Regression Model for default in terms of all the three predictors.

```{r}
model2 <- glm(default~balance+student+income, data = Default, family = binomial)
summary(model2)
```

## Question 2

### (a) Import the “Heart” dataset and view the first few rows. Investigate the structure/class of each variable.

```{r}
heartDS <- read.csv("../datasets/heart.csv")
attach(heartDS)
head(heartDS)
dim(heartDS)
str(heartDS)
```

By looking at the output we can see that heart data set consists of 303 observations of 15 variables.

We can see that our target variable "AHD" is the integer variable to we need to make it a factor variable because this a classification model.

```{r}
heartDS$AHD = as.factor(heartDS$AHD)
str(heartDS)
```

So now we can see that AHD is a factor variable now.


### (b) Fit a Logistic Regression Model for AHD in terms of Sex, ChestPain and Ca.

```{r}
model3 <- glm(AHD~Sex+ChestPain+Ca, data = heartDS, family = binomial)
summary(model3)
```


## Question 3

### (a) Load and attach the “Smarket” dataset and view the first few rows. Investigate the structure/class of each variable.

```{r}
attach(Smarket)
head(Smarket)
dim(Smarket)
str(Smarket)
```

Here we can see that, there are a total of 1250 observations of 9 variables.

Since, Direction is the categorical variable we can use it in our logistic regression model as a target variable.

### (b) Fit a Logistic Regression Model for Direction in terms of the Predictors Lag1, Lag2, Lag3, Lag4, Lag5 and Volume. 

```{r}
model4 <- glm(Direction~Lag1+Lag2+Lag3+Lag4+Lag5+Volume, data = Smarket, family = binomial)
summary(model4)
```

### (c) Explain how you would calculate the probability (the Direction = “up”) for a given set of X (predictor) values.

So to find the probability for Direction "Up", we need to find out that if our model has dummy variable for the same. To do so, we can use contrasts function

```{r}
contrasts(Direction)
```

Here, we can confirm that R has a dummy variable with 1 for "Up".

Now, we can check that coefficients of our model and create our model equation to find the probability of Direction "Up".

```{r}
coef(model4)
```

Now substituting the value of coefficients into our equation for the model,

$\hat{P(Direction = Up)}$ = $\frac{e^{\hat{-0.126000257} - \hat{0.073073746} Lag1 - \hat{0.042301344} Lag2 + \hat{0.011085108} Lag3 + \hat{0.009358938} Lag4 + \hat{0.010313068} Lag5 + \hat{0.135440659} Volume }}{1 + e^{\hat{-0.126000257} - \hat{0.073073746} Lag1 - \hat{0.042301344} Lag2 + \hat{0.011085108} Lag3 + \hat{0.009358938} Lag4 + \hat{0.010313068} Lag5 + \hat{0.135440659} Volume }}$

By using the above equation we can calculate the probability of Direction "Up".

### (d) Construct the Misclassification Martrix

for this we need to create a prediction class for the misclassification matrix first. For prediction class we first need to make predictions.

```{r}
predicted_prob <- predict(model4, type = "response")
```

From above we can see all the predictions of our model. Now we need to make a prediction class out of it for Up and Down. We will create a prediction class with the condition that if a probability is greater than 0.5 then we will take it as "Up".

```{r}
predicted_class <- rep("Down", nrow(Smarket))
predicted_class[predicted_prob > 0.5] = "Up"

table(predicted_class, Direction)
```

From the above table we can see that this is our misclassification matrix.

### (e) Calculate the Prob(Misclassification) or Misclassification Rate

Misclassification rate can be calculated as the number of items falsely predicted by our model prediction. From above table we can see that we falsely predicted **141** as **Down** and **457** as **Up**.

Therefore, using this misclassification rate would be:

```{r}
misclassificationRate <- (141 + 457) / (145 + 141 + 457 + 507)
misclassificationRate
```

From this we can find out that the misclassification rate is 47.84%.

### (f) Calculate the False Positive rate

From the misclassification matrix, false positive is 457. That is 457 items predicted as Up while in actual they were Down.

Therefore, using this false positive rate would be:

```{r}
falsePositiveRate <- 457 / (145 + 457)
falsePositiveRate
```

From above output, we can see that false positive rate is 75.91%

### (g) Calculate the False Negative rate

From the misclassification matrix, false negative is 141. That is 141 items predicted as Down while in actual they were Up

Therefore, using this false negative rate would be:

```{r}
falseNegativeRate <- 141 / (141 + 507)
falseNegativeRate
```

From above output, we can see that false negative rate is 21.75%